---
title: "RRproject1"
author: "lm"
date: "`r Sys.Date()`"
output: word_document
---

## Loading and preprocessing the data

```{r Q1}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,"repdata_data_activity.zip")
unzip("repdata_data_activity.zip")
data <- read.csv(unzip("repdata_data_activity.zip"))
```

## What is mean total number of steps taken per day?
The total number of steps taken per day was calculated and its histogram was plotted.
The mean and median of the total number of steps taken per day calculated and reported.

```{r Q2}
stepsperday <- tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(stepsperday, xlab = "Total step", main="Histogram of total step")
print(mean (stepsperday))
print(median (stepsperday))
```

## What is the average daily activity pattern?
a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r Q3}
xnames <- names(tapply(data$steps, data$interval, mean, na.rm = TRUE))
plot(tapply(data$steps, data$interval, mean, na.rm = TRUE),xaxt="n", type = "l", xlab= "interval", ylab = "steps")
axis(1, at=1:length(xnames), labels=xnames)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r Q4}
agrdata <- aggregate(steps~interval, data, mean, na.rm = TRUE)
max <- agrdata[which.max(agrdata$steps),]$interval
max
```
##Imputing missing values
the total number of missing values in the dataset (i.e. the total number of rows with NAs calculated.

A new dataset that is equal to the original dataset but with the missing data filled in was created.
```{r Q5}
missingValue <- sum(is.na(data$steps))
data2 <- data                                              
for(i in 1:ncol(data)) {                                   
data2[ , i][is.na(data2[ , i])] <- mean(data2[ , i], na.rm = TRUE)
}
```
A histogram of the total number of steps taken each day and  the mean and median total number of steps taken per day calculated. Imputing missing data on the estimates of the total daily number of steps creat a normal distribution according to the below historam and make mean and median closer to each other.
```{r Q6}
stepsperday2 <- tapply(data2$steps, data2$date, sum, na.rm = TRUE)
hist(stepsperday2, xlab = "Total step", main="Histogram of total step")
print(mean (stepsperday2))
print(median (stepsperday2))
```

```{r Q7}
data$date <- as.Date(strptime(data$date, format="%Y-%m-%d"))                      
weekdays <- weekdays(data$date)
data2 <- cbind(data,weekdays)
for (i in 1:nrow(data2)) {
        if (data2$weekdays[[i]] %in% c("Saturday","Sunday")) {
                data2$weekdays[[i]]<-"weekend"
        }
        else{
                data2$weekdays[[i]]<-"weekday"
        }

}
stepsByDay <- aggregate(data2$steps ~ data2$interval + data2$weekdays, data2, mean)
names(stepsByDay) <- c("interval", "weekdays", "steps")
library(ggplot2)
par(mfrow=c(1,2))
ggplot(stepsByDay, aes(interval, steps))+
        facet_grid(~weekdays)+
geom_line()

```
##Are there differences in activity patterns between weekdays and weekends?